Computing in the Field of Rationals

### 2012-02-27 rational.tex
_

\begin{abstract}
  Every functional programmer knows about sum and product types, $a+b$
  and $a*b$ respectively. Negative and fractional types, $a-b$ and
  $a/b$ respectively, are much less known and their computational
  interpretation is unfamiliar and often complicated. We show that in
  a programming model in which information is preserved (such as the
  model introduced in our recent paper on \emph{Information Effects}),
  these types have particularly intuitive and natural computational
  interpretations. Intuitively, values of negative types are values
  that flow ``backwards'' to satisfy demands, values of fractional
  types are values that represent first-class ``structural''
  constraints.
  %% , and values of square root and imaginary types are values that can
  %% be related by arbitrary algebraic constraints.
  The combination of these negative, fractional, square root, and
  imaginary types enable greater flexibility in programming by
  breaking complicated invariants into local ones that can be
  independently satisfied by a subcomputation. Furthermore, they allow
  a programmer to express a wide range of programming idioms,
  including higher-order functions, delimited continuations,
  speculative computation, logic programming, and more.
\end{abstract}

>>>>

\begin{abstract}
  Every functional programmer knows about sum and product types,
  {{a+b}} and {{a*b}} respectively. Negative and fractional types,
  $a-b$ and $a/b$ respectively, are much less known and their
  computational interpretation is unfamiliar and often complicated. We
  show that in a programming model in which information is preserved
  (such as the model introduced in our recent paper on
  \emph{Information Effects}), these types have particularly intuitive
  and natural computational interpretations. Intuitively, values of
  negative types are values that flow ``backwards'' to satisfy
  demands, values of fractional types are values that represent
  first-class ``structural'' constraints.  The combination of these
  negative and fractional types enable greater flexibility in
  programming by breaking complicated invariants into local ones that
  can be independently satisfied by a subcomputation. Furthermore,
  they allow a programmer to express a wide range of programming
  idioms, including higher-order functions, delimited continuations,
  speculative computation, logic programming, and more.

\end{abstract}
